\SuperScreen: An open-source package for simulating the magnetic response of two-dimensional superconducting devices

### 2022-06-06 main.tex

Here, we briefly describe existing software tools for modeling the magnetic response of superconducting devices, most of which are specifically designed for inductance extraction for superconducting integrated circuits. FastHenry, a widely-used 3D (normal metal) inductance extraction tool from MIT~\cite{Kamon1994-ck}, has been extended to support superconducting elements~\cite{wrcad, XicTools}. This modified version of FastHenry has been used for inductance extraction in the commercial software InductEx~\cite{Fourie2011-wl}. While FastHenry is open-source, it is written in C and must be compiled for a specific computer architecture and operating system. FastHenry executables compiled for several common operating systems are available as part of the open-source XicTools superconducting integrated circuit design suite from Whiteley Research, Inc~\cite{wrcad, XicTools}. A 2D London model based on a scalar stream function, much like the model used by \SuperScreen~\cite{Brandt2004-ew, Brandt2005-wj}, forms the basis of the 3D-MLSI software package~\cite{Khapaev1997-kw, Khapaev2001-xq, Khapaev2001-pw}, which is also written in C and is not open-source. For a more thorough overview and comparison of inductance extraction tools, see Ref.~\cite{Gaj1999-ls}.

@@

If one defines an inhomogeneous effective penetration depth $\Lambda(x, y)$ in a \SuperScreen model, Equation~\ref{eq:london_inhomogeneous_2d}, rather than Equation~\ref{eq:london}, is solved numerically as follows. For a mesh with $p$ vertices, the effective penetration depth is represented by a $p\times 1$ vector $\mathbf{\Lambda}$. Equations~\ref{eq:Heff_sub_num} and \ref{eq:full_stream} are updated according to:
\begin{align}
\begin{split}
    \label{eq:Heff_inhomogeneous}
    &\mathbf{Q}.\mathbf{w}^T-\Lambda\mathbf{\nabla}^2\to\\
    &\mathbf{Q}.\mathbf{w}^T-\mathbf{\Lambda}^T.\mathbf{\nabla}^2-\vec{\nabla}\mathbf{\Lambda}\cdot\vec{\nabla}
\end{split}
\end{align}
The notation $\vec{\nabla}\mathbf{f}\cdot\vec{\nabla}$ indicates an inner (dot) product over the two spatial dimensions, resulting in a $p\times p$ matrix such that $(\vec{\nabla}\mathbf{f}\cdot\vec{\nabla})\mathbf{g}$ computes $(\vec{\nabla}f(x, y))\cdot(\vec{\nabla}g(x, y))$ (see~\ref{appendix:laplace}).

>>>>

Here, we briefly describe existing software tools for modeling the magnetic response of superconducting devices, most of which are specifically designed for inductance extraction for superconducting integrated circuits. FastHenry, a widely-used 3D (normal metal) inductance extraction tool from MIT~\cite{Kamon1994-ck}, has been extended to support superconducting elements~\cite{wrcad, XicTools}. This modified version of FastHenry has been used for inductance extraction in the commercial software InductEx~\cite{Fourie2011-wl, Fourie2012-gv, Jackman2016-mf}. While FastHenry is open-source, it is written in C and must be compiled for a specific computer architecture and operating system. FastHenry executables compiled for several common operating systems are available as part of the open-source XicTools superconducting integrated circuit design suite from Whiteley Research, Inc~\cite{wrcad, XicTools}. A 2D London model based on a scalar stream function, much like the model used by \SuperScreen~\cite{Brandt2004-ew, Brandt2005-wj}, forms the basis of the 3D-MLSI software package~\cite{Khapaev1997-kw, Khapaev2001-xq, Khapaev2001-pw, Khapaev2002-ev}, which is also written in C and is not open-source. For a thorough historical overview and comparison of inductance extraction tools, see Ref.~\cite{Gaj1999-ls}. For a more recent overview of superconducting inductance extraction tools see Ref.~\cite{Tolpygo2021-jz}, particularly references [8--25] therein.

\SuperScreen is not intended primarily as an inductance extraction tool and cannot be used to extract the inductance of superconducting transmission lines (a common superconducting circuit design task~\cite{Gaj1999-ls,Khapaev1996-uu,Tolpygo2021-jz}) due to the lack of support for ``terminal currents.'' Nevertheless, it is instructive to compare \SuperScreen to a state of the art commercial inductance extraction tool, InductEx~\cite{Fourie2011-wl, Fourie2012-gv, Jackman2016-mf}, on a problem that can be solved using both tools. One such problem, provided as an example in the InductEx user manual, is the self-inductance $L$ of a square washer with outer dimension $30\,\um$, inner dimension $10\,\um$, London penetration depth $\lambda=0.24\,\um$, and film thickness $d=0.20\,\um$, for which the InductEx solution reports a ``design'' value of $L=20\,\mathrm{pH}$ and an ``extracted'' value of $L=20.0956\,\mathrm{pH}$~\cite{inductex_user_manual_2020}. Code Block~\ref{code:washer} demonstrates how to solve this problem using \SuperScreen. The inductance extracted using \SuperScreen is within 1\% of the inductance extracted using InductEx.

\begin{code-onecol}
\begin{minted}[fontsize=\small]{python}
import superscreen as sc
from superscreen.geometry import box

layer = sc.Layer(
    "base",
    london_lambda=0.24,
    thickness=0.2,
)
film = sc.Polygon(
    "washer",
    layer="base",
    points=box(30, points_per_side=50),
)
hole = sc.Polygon(
    "hole",
    layer="base",
    points=box(10, points_per_side=250),
)
device = sc.Device(
    "washer",
    layers=[layer],
    films=[film],
    holes=[hole],
    length_units="um",
)
device.make_mesh(min_points=3000)
# See Equation 19.
M = device.mutual_inductance_matrix()
L = M[0, 0].to("pH")
print(f"Inductance = {L:~.4fP}")
# Output: Inductance = 19.9100 pH
\end{minted}
\captionof{listing}{\SuperScreen script to calculate the self-inductance of a square washer. The extracted inductance $L=19.91\,\mathrm{pH}$ differs from the InductEx ``design'' and ``extracted'' values by -0.45\% and -0.92\% respectively~\cite{inductex_user_manual_2020}. }
\label{code:washer}
\end{code-onecol}

@@

If one defines an inhomogeneous effective penetration depth $\Lambda(x, y)$ in a \SuperScreen model, Equation~\ref{eq:london_inhomogeneous_2d}, rather than Equation~\ref{eq:london}, is solved numerically as follows. For a mesh with $p$ vertices, the effective penetration depth is represented by a $p\times 1$ vector $\mathbf{\Lambda}$. Equations~\ref{eq:Heff_sub_num} and \ref{eq:full_stream} are updated according to:
\begin{align}
\begin{split}
    \label{eq:Heff_inhomogeneous}
    &\mathbf{Q}.\mathbf{w}^T-\Lambda\mathbf{\nabla}^2\to\\
    &\mathbf{Q}.\mathbf{w}^T-\mathbf{\Lambda}^T.\mathbf{\nabla}^2-\vec{\nabla}\mathbf{\Lambda}\cdot\vec{\nabla}
\end{split}
\end{align}

The notation $\vec{\nabla}\mathbf{f}\cdot\vec{\nabla}$ indicates an inner (dot) product over the two spatial dimensions, resulting in a $p\times p$ matrix such that $(\vec{\nabla}\mathbf{f}\cdot\vec{\nabla})\mathbf{g}$ computes $(\vec{\nabla}f(x, y))\cdot(\vec{\nabla}g(x, y))$ (see~\ref{appendix:laplace}).
